A sheaf--function dictionary for algebraic tori over local fields

### 2013-08-18 shanghai.tex
Added the proof of the dictionary for algebraic tori.

In Part 2 we use Theorem~\ref{thm:snake} to make a sheaf--function dictionary
for algebraic tori over local fields.  Let $K$ be a non-archimedian local field
with finite residue field $\Fq$; we place no restriction on the characteristic of $K$.
Write $\OK$ for the ring of integers of $K$ and $\pK$ for the maximal ideal in $\OK$.
Let $T$ be a torus defined over $K$.  The N\'eron model $\TT$ of $T$
is a smooth group scheme over $\OK$ with generic fiber $T$ and a canonical isomorphism
$\TT(\OK) \cong T(K)$.  For each positive integer $d$, let $\GN{T}_d$ be the
Greenberg transform of $\TT \times_{\Spec{\OK}} \Spec{\OK/\pK^d}$; the commutative
group scheme $\GN{T}_d$ plays the r\^{o}le of $G$.  Theorem~\ref{thm:application}
allows us to translate characters $T(K)$ of depth less than $d$ into quasicharacter sheaves on $\GN{T}_d$:
\[
\QCiso{\GN{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx).
\]  
We refer to finite-depth characters of $T(K)$ as quasicharacters and
describe a group scheme $\GN{T}$ simultaneously geometrizing all such quasicharacters.

@@

For any $n\leq m$ (positive integers) define $f_{n\leq m} : \GN{T}_{m} \to \GN{T}_{n}$ by $f_{n\leq m}\ceq f_n \circ f_{n+1} \circ \cdots \circ f_{m-1}$.
Let $f_{n\leq m}^* : \QC(\GN{T}_n) \to \QC(\GN{T}_{m})$ be the functor promised by Proposition~\ref{prop:pullback}.

\begin{proposition}
The functor $f_{n\leq m}^* : \QC(\GN{T}_n) \to \QC(\GN{T}_{m})$ is full and faithful.
\end{proposition}

\begin{proof}
It suffices to find a morphism of schemes $i_n : \GN{T}_{n}\hookrightarrow \GN{T}_{n+1}$ such that $f_n\circ i_n = \id_{\GN{T}_n}$. Base change along $\OK/\pK^{n+1} \to \OK/\pK^n$ determines a closed immersion of schemes $\TT_{n+1} \hookrightarrow \TT_n$ (not a morphism of group schemes). Applying the Greenberg functor produces a closed immersion of schemes $i_n : \GN{T}_{n} \hookrightarrow \GN{T}_{n+1}$. 
\Clifton{I still need to confirm this claim.}
Although $i_n$ is not a morphism of group schemes, it is a section of $f_n : \GN{T}_{n+1}\to \GN{T}_n$.
\end{proof}

We now define a category, denoted by $\QC(\GN{T})$, which allows us to extend the notion of quasicharacter sheaf from $\GN{T}_n$ to $\GN{T}$. An object in $\QC(\GN{T})$ is an equivalence class in $\cup_{n} \obj \QC(\GN{T}_n)$ under the relation: $\qcs{L}_n\in \QC(\GN{T}_n)$ is equivalent to $\qcs{L}_{m}\in \QC(\GN{T}_m)$ if $f_{n\leq k}^* \qcs{L}_{n} = f_{m\leq k}^* \qcs{L}_m$ for some $n,m\leq k$. Objects in $\QC(\GN{T})$ all take the form
\[
[\qcs{L}_n] \ceq \{ f_{n\leq k}^* \qcs{L}_n \tq n\leq k\},
\]
for some unique $\qcs{L}_n \in \QC(\GN{T}_n)$.
Maps $[\qcs{L}_n] \to [\qcs{L}_m]$ in $\QC(\GN{T})$ are defined as follows:
$\Hom([\qcs{L}_n],[\qcs{L}_m])$ is the set of equivalence classes in $\cup_{n,m\leq k} \Hom(f_{n\leq k}^*\qcs{L}_n,f_{m\leq k}^*\qcs{L}_m)$ under the relation: $\alpha_k\in \mor \QC(\GN{T}_k)$ is equivalent to $\beta_l \in \QC(\GN{T}_l)$ if $f_{k\leq i}^*\alpha_k = f_{l\leq i}^* \beta_l$ for some $k,l\leq i$. Maps in $\QC(\GN{T})$ all take the form
\[
[\alpha_k] \ceq \{ f_{k\leq i}^* \alpha_k \tq k\leq i\},
\]
for some unique $\alpha_k \in \QC(\GN{T}_k)$.
Identities and composition are defined in the natural way.
Category $\QC(\GN{T})$ carries a tensor structure under
\[
[\qcs{L}_n]\otimes [\qcs{L}_m] \ceq [f_{n\leq k}^*\qcs{L}_n \otimes f_{m\leq k}^*\qcs{L}_m],
\qquad k = \max\{ n,m\}.
\]
Duals in $\QC(\GN{T})$ are defined by
\[
[\qcs{L}_n]^\vee \ceq [\qcs{L}_n^\vee].
\]
Category $\QC(\GN{T})$ is rigid symmetric monoidal and comes equipped with monoidal subcategories
\[
\begin{tikzcd}
\ & & & \QC(\GN{T})\\
\arrow{urrr}{F_n} \QC(\GN{T}_n) \arrow[swap]{rr}{f_{n\leq m}^*} && \QC(\GN{T}_m) \arrow[swap]{ur}{F_m} & \\
\end{tikzcd}
\]
defined by $F_n(\qcs{L}) = [\qcs{L}_n]$ and $F_n(\alpha_n) = [\alpha_n]$, for all positive integers $n\leq m$.
We refer to objects of $\QC(\GN{T})$ as \emph{quasicharacter sheaves on $\GN{T}$} or \emph{quasicharacter sheaves for $T$}.

@@

\begin{proof}
Show $\QCiso{\GN{T}} = \varinjlim_{d\in \ZZ_{>0}} \QCiso{\GN{T}_d}$, then invoke Theorem~\ref{thm:application}.
\end{proof}

>>>>

In Part 2 we use Theorem~\ref{thm:snake} to make a sheaf--function dictionary
for algebraic tori over local fields.  Let $K$ be a non-archimedian local field
with finite residue field $\Fq$; we place no restriction on the characteristic of $K$.
Write $\OK$ for the ring of integers of $K$ and $\pK$ for the maximal ideal in $\OK$.
Let $T$ be a torus defined over $K$.  The N\'eron model $\TT$ of $T$
is a smooth group scheme over $\OK$ with generic fiber $T$ and a canonical isomorphism
$\TT(\OK) \cong T(K)$.  For each positive integer $d$, let $\GN{T}_n$ be the
Greenberg transform of $\TT \times_{\Spec{\OK}} \Spec{\OK/\pK^n}$; the commutative
group scheme $\GN{T}_d$ plays the r\^{o}le of $G$.  Theorem~\ref{thm:application}
allows us to translate characters $T(K)$ of depth less than $d$ into quasicharacter sheaves on $\GN{T}_d$:
\[
\QCiso{\GN{T}_n} 
\iso
\Hom_{< n}(T(K),\EEx).
\]  
We refer to finite-depth characters of $T(K)$ as quasicharacters and
describe a group scheme $\GN{T}$ simultaneously geometrizing all such quasicharacters.

@@

For any $n< m$ (positive integers) define $f_{n\leq m} : \GN{T}_{m} \to \GN{T}_{n}$ by $f_{n\leq m}\ceq f_n \circ f_{n+1} \circ \cdots \circ f_{m-1}$; also set $f_{n\leq n} \ceq \id_{\GN{T}_n}$.
Let $f_{n\leq m}^* : \QC(\GN{T}_n) \to \QC(\GN{T}_{m})$ be the functor promised by Proposition~\ref{prop:pullback}.
%
%\begin{proposition}\label{prop:nm}
The functor $f_{n\leq m}^* : \QC(\GN{T}_n) \to \QC(\GN{T}_{m})$ is full and faithful.
%\end{proposition}
%
%\begin{proof}
%It suffices to find a morphism of schemes $i_n : \GN{T}_{n}\hookrightarrow \GN{T}_{n+1}$ such that $f_n\circ i_n = \id_{\GN{T}_n}$. Base change along $\OK/\pK^{n+1} \to \OK/\pK^n$ determines a closed immersion of schemes $\TT_{n+1} \hookrightarrow \TT_n$ (not a morphism of group schemes). Applying the Greenberg functor produces a closed immersion of schemes $i_n : \GN{T}_{n} \hookrightarrow \GN{T}_{n+1}$. 
%\Clifton{I still need to confirm this claim.}
%Although $i_n$ is not a morphism of group schemes, it is a section of $f_n : \GN{T}_{n+1}\to \GN{T}_n$.
%\end{proof}

We now define a category which allows us to extend the notion of quasicharacter sheaf to the pro-group scheme $\GN{T}$.

\begin{definition}
An object in $\QC(\GN{T})$ is an equivalence class in $\cup_{n} \obj \QC(\GN{T}_n)$ under the relation: $\qcs{L}_n\in \QC(\GN{T}_n)$ is equivalent to $\qcs{L}_{m}\in \QC(\GN{T}_m)$ if $f_{n\leq k}^* \qcs{L}_{n} = f_{m\leq k}^* \qcs{L}_m$ for some $n,m\leq k$. Objects in $\QC(\GN{T})$ all take the form
\[
[\qcs{L}_n] \ceq \{ f_{n\leq k}^* \qcs{L}_n \tq n\leq k\},
\]
for some unique $\qcs{L}_n \in \QC(\GN{T}_n)$.
Maps $[\qcs{L}_n] \to [\qcs{L}_m]$ in $\QC(\GN{T})$ are defined as follows:
$\Hom([\qcs{L}_n],[\qcs{L}_m])$ is the set of equivalence classes in $\cup_{n,m\leq k} \Hom(f_{n\leq k}^*\qcs{L}_n,f_{m\leq k}^*\qcs{L}_m)$ under the relation: $\alpha_k\in \mor \QC(\GN{T}_k)$ is equivalent to $\beta_l \in \QC(\GN{T}_l)$ if $f_{k\leq i}^*\alpha_k = f_{l\leq i}^* \beta_l$ for some $k,l\leq i$. Maps in $\QC(\GN{T})$ all take the form
\[
[\alpha_k] \ceq \{ f_{k\leq i}^* \alpha_k \tq k\leq i\},
\]
for some unique $\alpha_k \in \QC(\GN{T}_k)$.
Identities and composition are defined in the natural way.
We refer to objects in $\QC(\GN{T})$ as \emph{quasicharacter sheaves on $\GN{T}$} or \emph{quasicharacter sheaves for $T$}.
\end{definition}

Category $\QC(\GN{T})$ carries a tensor structure under
\[
[\qcs{L}_n]\otimes [\qcs{L}_m] \ceq [f_{n\leq k}^*\qcs{L}_n \otimes f_{m\leq k}^*\qcs{L}_m],
\qquad k = \max\{ n,m\}.
\]
Duals in $\QC(\GN{T})$ are defined by $[\qcs{L}_n]^\vee \ceq [\qcs{L}_n^\vee]$.

@@

\begin{proof}
Observe that $\QC(\GN{T})$ comes equipped with monoidal subcategories
\[
\begin{tikzcd}
\ & & & \QC(\GN{T})\\
\arrow{urrr}{F_n} \QC(\GN{T}_n) \arrow[swap]{rr}{f_{n\leq m}^*} && \QC(\GN{T}_m) \arrow[swap]{ur}{F_m} & \\
\end{tikzcd}
\]
defined by $F_n(\qcs{L}) = [\qcs{L}_n]$ and $F_n(\alpha_n) = [\alpha_n]$, for all positive integers $n\leq m$.
Passing to isomorphism classes of objects, we get 
\[
\begin{tikzcd}
\ & & & & \QCiso{\GN{T}}\\
\ & & & \varinjlim_{d\in \ZZ_{>0}} \QCiso{\GN{T}_d} \arrow{ur}{\exists ! \theta} & \\
\arrow{urrr}{F_n} \QCiso{\GN{T}_n} \arrow[swap]{rr}{f_{n\leq m}^*} && \QCiso{\GN{T}_m} \arrow[swap]{ur}{F_m} & & \\
\end{tikzcd}
\]
The group homomorphism 
$\theta: \varinjlim_{d\in \ZZ_{>0}} \QCiso{\GN{T}_d} \to \QCiso{\GN{T}}$ is a bijection 
with inverse $[\qcs{L}]_{/\text{iso}} \mapsto [\qcs{L}_{/\text{iso}}]$.
Now use Theorem~\ref{thm:application}.
\end{proof}
